Casual Bandits

### 2015-12-24 causal-bandit/icml-2015/suplementary_proofs.tex
begins paper outline

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{comment}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{tikz} 
\usetikzlibrary{arrows,positioning} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left\{#1\right\}}

@@

\begin{comment}
\section{Random notes}
Information directed sampling have developed a general framework to construct algorithms that leverege generic structure between arms to minimize Bayesian regret, by explicitly trading off expected regret and information gain at each timestep.

What can be expressed in the partial monitoring framework that cannot be defined by the feedback graph?

We consider a variant of the stochastic multi-armed bandit problem when we have prior knowledge of the causal structure, but not the functional form, of the relationship between variables.

sequential decision process,exploration-exploitation trade off

Points to present. Why we use regret against single best arm - it makes sense in the stochastic setting. 
The difference between observational and interventional distributions.

terminology for intervention vs observation. Recast bandit algorithm in terms of this. Show limitations to existing approaches by pointing to places were existing algorithms eg UCB/explore-exploit/exp3 require unbiased estimates. Are unbiased estimates actually impossible in our setting?

This paper is about sequentially choosing actions. 
Modelling sequntial decition processes.

In the classic multi-armed bandit setting, an ag

General bandit part ... define multi-armed bandit, and importance. Decision making is fundamental. Goal is to select an action to maximize a reward (or minimize a loss). In many settings, such as drug trials, online ad placement, etc we can choose actions sequentially. This introduces an exploration-exploitation trade-off.

The problem of sequentially choosing actions to minimize some loss wide range of applications in XXXX, YYYY. The multi-armed bandit setting, in which .... .

Now the required causal stuff should go here.

Causal structure can be represented as a directed acyclic graph with nodes representing the variables and an edge from $A \rightarrow B$ meaning $A$ causes $B$. Deterministic interventions, consisting of setting variables to some value, are modelled by setting the nodes to the desired values and removing any links coming into them, as their value is now no longer set by other variables in the graph according to the original mechanism.

If the causal graph is known, such that there are no hidden variables causing 2 or more of the variables and the probabilities are all > 0 then the distribution of any variable after any intervention can be inferred (asymptotically) from data collected in absence of any interventions.
\end{comment}

\section*{Introduction}

Problems requiring choosing an action under uncertainty are rife in all areas of human endeavour. For many problems, actions may be chosen sequentially, allowing the agent to learn from the outcome of early choices to improve later ones. A widely used framework for sequential decision making is the multi-armed bandit. In the classic multi-armed bandit setting there is a finite set of available actions, each associated with a distribution over rewards which is unknown but stationary and independent of the reward distribution of other actions. At each timestep the agent selects an action and receives a reward sampled iid from the corresponding reward distribution.

An an alternate approach to selecting actions is causal inference. Frameworks for causal inference provide a mechanism to specify assumptions that allow observational distributions over variables to be mapped to interventional ones. This allows an agent to predict the outcome of an action based on non-experimental data. This approach is common in social science, demography, and economics where explicit experimentation may be difficult. For example, predicting the effect of changes to childcare subsidies on workforce participation or school choice on student grades.

We take a first step towards unifying these approaches by considering a variant of the stochastic multi-armed bandit problem where we have prior knowledge of the causal structure governing the available actions. This structure creates dependencies between the rewards of different arms such that selecting one action can provide information on the reward for other actions.

There has been substantial recent work into extending bandit algorithms to incorporate additional assumptions and deal with more complex feedback structures. Algorithms with strong guarantees have been developed for linear bandits [], generalized linear bandits, gaussian process bandits [], etc. There is also an active line of research into bandits with feedback defined by a graph. Actions are modelled as nodes in the graph and the agent observes rewards for each action connected to the selected action []. The novelty of our work is that we assume prior knowledge of the causal structure but not the functional form of the relationship between variables.

\section*{Problem Formulation}
Assume we have a known causal model with binary variables $\boldsymbol{X} = \{X_{1}..X_{N}\}$ that independently cause a target variable of interest $Y$, figure \ref{fig:causalStructure}. We can run sequential experiments on the system, where at each timestep $t$ we can select a variable on which to intervene and subsequently observe the complete result, $(\boldsymbol{X}_{t},Y_{t})$. As an example, consider a farmer wishing to optimize the yield of her crop. She can invest in a green house to control temperature, a watering system to control soil moisture, fertilizers to set soil nutrients, etc. We assume only a single intervention is feasible due to cost and that each of these variables are independent of one-another (this may not always be the case - temperature could be related to rainfall for example). After having selected which variable to control, she plants her crops and observes the values of the remaining input variables and the yield. This repeats across many growing seasons, and the goal is to maximize the total cumulative yield.

\begin{figure}[h]
\centering
\caption{Assumed Causal Structure}
\label{fig:causalStructure}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed}, hidden/.style={empty}]

 %nodes
\node[main node](1){$X_{1}$};
\node[main node, right=of 1](2){$X_{2}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){Y};
 \path[every node/.style={font=\sffamily\small}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5);

\end{tikzpicture}
\end{figure}

Let $q \in [0,1]^N$ be a fixed vector where $q_i = P(X_i = 1)$. In each time-step $t$ upto a known end point $T$:

\begin{enumerate}
\item The learner chooses an $I_t \in \set{1,\ldots, N}$ and $J_t \in \set{0,1}$.
\item Then $X_t \in \set{0,1}^N$ is sampled from a product of Bernoulli distributions, $X_{t,i} \sim \Ber(q_i)$ 
\item The learner observes $\tilde X_t \in \set{0,1}^K$, which is defined by 
\eq{
\tilde X_{t,i} = \begin{cases}
X_{t,i} &\text{if } i \neq I_t \\
J_t & \text{otherwise}\,.
\end{cases}
}
\item The learner receives reward $Y_t \sim \Ber(r(\tilde X))$ where $r:\set{0,1}^K \to [0,1]$ is unknown and arbitrary. THIS LOOKS WRONG SHOULD BE N not K.
\end{enumerate}

The expected reward of taking action $i,j$ is $\mu_{i,j} = \E[r(X)|do(X_i = j)]$. The optimal reward and action are denoted $\mu^*$ and $(i^*,j^*)$ respectively,
where $(i^*,j^*) = \argmax_{i,j} \mu_{i,j}$ and $\mu^* = \mu(i^*,j^*)$. The $n$-step cumulative expected regret is
\eq{
R_n = \E \sum_{t=1}^n \left(\mu^* - \mu_{I_t,J_t}\right).
}

The problem can be treated as a classical multi-armed bandit with $K = 2N$ arms. However, this does not utilize the information provided by the causal assumption.

Now need to expand upon how the assumption gives us extra information.

The difficulties faced due to bias.

To deal with this issue we use a simple explore-exploit algorithm. Our algorithm will explore for $h$ time-steps, sampling actions in a way that depends on our prior knowledge of $q$ but is independent of the observed rewards. We then select the arm with the highest expected reward for the remaining $T-h$ time steps.

\section*{Results}

Summarize results here and note differences to classic bandit results

\subsection*{Known and Balanced q}

We begin with the simplest case were we assume that $q_i=\frac{1}{2} \forall i$. During the exploration phase we sample actions uniformly at random. In this case, this is equivalent to purely observing, that is taking no action and allowing all input variables to take their value randomly as  $X_{t,i} \sim \Ber(\frac{1}{2})$.

We will have $n_i \sim Binomial(h,\frac{1}{2})$ observations for each arm $i$ at the end of the exploration stage. Note that this is independent of the number of arms $K$.

Assume we have $K$ bernoulli arms with means ordered from highest to lowest $\mu_1 ... \mu_K$. Let $\Delta = [\Delta_1...\Delta_K]$ be the differences from the optimal reward $\mu_1$.

\subsubsection*{Regret during explore phase}
Since the probability we play each arm is constant and uniform during the exploration phase, the expected regret is simply proportional to the average sub-optimality $\Delta$.
\eqn{
R_1 = h\sum_i P(i)\Delta_i = \frac{h}{K}\sum_i \Delta_i = h E[\Delta] < h
}

\subsubsection*{Regret during exploit phase}
The regret during this phase is proportional to the expected sub-optimality of the arm with the highest empirical mean at the end of the explore phase.

\eqn{
\hat{i^*} = argmax_i [\hat{\mu}_i]
}
\eqn{
R_2 = (T-h)E[\Delta_{\hat{i^*}}] = (T-h)\sum_i P(\hat{\mu}_i \geq \hat{\mu}_j \forall j)\Delta_i 
}

The difficulty with this approach is that it is hard to get bounds that are tight for all $\Delta$. Instead, we will bound the probability that we select an arm with a sub-optimality gap greater than some $D$.
\eqn{
R_2 \leq (T-h)\left(P(\Delta_{\hat{i^*}} \leq D) D+P(\Delta_{\hat{i^*}} > D)  \Delta_{max} \right)
}
The goal now is to get a bound for $P(\Delta_{\hat{i^*}} > D)$ in terms of Hoeffdings type bounds for each arm.

Suppose $i = \hat{i^*} \implies \hat{\mu}_i > \hat{\mu}_1$. If we haven't over-estimated $\mu_i$ too much, $\hat{\mu}_i - \mu_{i} < \frac{D}{2}$, and haven't under-estimated $\mu_1$ too much, $\mu_1 - \hat{\mu_1} < \frac{D}{2}$, then $\Delta_{\hat{i^*}} = \mu_1 - \mu_i < D$

\eqn{
\label{eqn:probDeltaToLarge}
P(\Delta_{\hat{i^*}} > D) \leq  P(\mu_1 - \hat{\mu_1} > \frac{D}{2})+ \sum_{i=2}^K P(\hat{\mu}_i - \mu_{i} > \frac{D}{2})
}

If we used the empirical mean as an estimator for $\mu_i$, the bound will depend on the number of times we actually observed each arm, which will be a random variable drawn from a multinomial distribution. Instead we will use an importance weighted estimator.

\eqn{
\label{eqn:importance_weighted_estimator}
\hat{\mu}_i = \frac{1}{h}\sum_{t=1}^h \frac{Y_t\ind{\text{arm $i$ active}}}{q_i}
}

where $q_i = P(\text{arm $i$ active})$

Hoeffdings gives $ P(\hat{\mu}_i - \mu_{i} > \epsilon) \leq e^{-2h\epsilon^2q_i^2}$. In this case we have assumed $q_i = \frac{1}{2} \forall i$. Putting this into equation \ref{eqn:probDeltaToLarge}:

\eqn{
\label{eqn:balancedHoeffdings}
P(\Delta_{\hat{i^*}} > D) \leq Ke^{-hD^2/8}
}

\eqn{
R_2 \leq (T-h)[(1-K e^{-hD^2/8})D + K e^{-hD^2/8}] < (T-h)[D + K e^{-hD^2/8}]
}

Let $D = \sqrt{\frac{8}{h}\log(hK)}$

\eqn{
R_2 \leq (T-h)\left(\sqrt{\frac{8}{h}\log(hK)} + \frac{1}{h}\right)
}

\subsection*{Total Regret}

Putting together the regret from the exploration and exploitation phases,

\eqn{
R_T & \leq \frac{h}{K}\sum_i \Delta_i + (T-h)\left(\sqrt{\frac{8}{h}\log(hK)} + \frac{1}{h}\right)\\
& \leq h + T\left(\sqrt{\frac{8}{h}\log(TK)} + \frac{1}{h}\right)
}

Now if we let $h = T^{2/3}(\log(KT))^{1/3}$,

\eqn {
R_T \leq 4T^{\frac{2}{3}}(log(KT))^{\frac{1}{3}} + T^{\frac{1}{3}}(log(KT))^{-\frac{1}{3}}
}

If $T \geq 2$ and $K \geq 2$, the first term dominates and,

\eqn {
R_T  \leq 5T^{\frac{2}{3}}(log(KT))^{\frac{1}{3}}
}

The distribution independent lower bound for optimised UCB is $O(\sqrt{TK})$ (see Bubeck sect 2.4.3) so we would expect our algorithm to do better if $K >> T^{\frac{1}{3}}$

\subsection*{Empirical results}
\begin{figure}[H]
\centering
\caption{Comparison of the UCB and causal-explore-exploit for K=20 and T=10000. Note, $K \sim T^{1/3}$ Plot shows average and standard deviation over 10000 trials.}
\includegraphics[width=.5\textwidth]{explore_exploit}
\end{figure}

\section{Generalizing to unbalanced $\boldsymbol{q}$}

When some arms have low natural probability we cannot rely on exploring them adequately by pure observation. We need to explicitly play them during the exploration phase.

We now have an additional trade off to make, which is how much should be observe (learning something about half the arms each timestep) versus playing the low probability arms.

Without loss of generality, we can assume $q_i \in [0,\frac{1}{2}]$ and $q_1 \leq q_2 ... \leq q_N$. Let $m \in [2,N] = \{m:q_m > \frac{1}{m}\}$ Ie if the problem is completely balanced $q_1...q_N = \frac{1}{2}$ then $m = 2$. If the problem is completely unbalanced, $q_1...q_N = 0$ then $m=N$. Let $q_{ij} = \P{X_i = j}$

Suppose we observe for the first $h/2$ timesteps. This is at worst half the optimal.

We then have estimates

\eqn {
\hat{\mu}_{ij} = \frac{\sum_{t=1}^{h/2}\ind{Y_t=1,X_{t,i}=j}}{\frac{h}{2}q_{ij}}
}

We take this as our estimate for those arms for which $q_{ij} > \frac{1}{m}$

@@

We play each of the remaining $m$ arms $\frac{h}{2m}$ times so for them we get

@@

Again comparing to the UCB bound of $\sqrt{KT}$, we expect the causal algorithm to do better if $K > m^{2/3}T^{1/3}$, or alternatively if $m \leq \frac{K^{3/2}}{\sqrt{T}}$

To get results that degrade to similar order bounds as UCB when the arms are very unbalanced, I will need to drop the explore/exploit strategy.

\subsection{Generalizing to unknown q}
We now consider the case where $\boldsymbol{q}$, and thus $m$ is not known in advance. We begin by considering the simple regret

\subsubsection{Bounding the Simple Regret}

\begin{itemize}
\item $N$ is number of variables.
\item $q_i$ is probability that $X_{i,t} = 1$ for any $t$
\item $s_i = \min\set{q_i, 1 - q_i}$
\item $\mu : \set{1,\ldots,N} \times \set{0,1} \to [0,1]$.
\item $I_t \in \set{1,\ldots,N} \times \set{0,1}$ is the intervention
\end{itemize}

Let $R_s(h)$ be the simple regret after $h$ time steps, which is
\eq{
R_s(h) = \E\left[\mu^* - \mu_{I_h}\right]\,,
}
where $\mu^*$ is the mean payoff of the optimal intervention and $I_h$ is the intervention chosen in the $h$th round.
We propose a two part algorithm that is provably order optimal when its performance is measured in terms of the simple regret.

\begin{algorithm}[H]
\caption{Simple Regret Algorithm}\label{alg:simple}
\begin{algorithmic}[1]
\State {\bf Input:} $h, N$
\For{$t \in 1,\ldots,h/2$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute for all $i \in \set{1,\ldots,N}$ and $j \in \set{0,1}$:
\eq{
\hat \mu_{i,j} = \frac{\sum_{t=1}^{h/2} \ind{X_{i,t} = j} r_t}{\sum_{t=1}^{h/2} \ind{X_{i,t} = j}}\,.
}
\State Compute $\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} X_{i,t}$
\State Compute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Compute $\hat{s'} = sorted(\hat{s}) : \hat{s'}_1 \leq \hat{s'_2} \leq ... \leq \hat{s'}_N$
\State Compute $\hat m = \min\set{2 \leq i \leq N : \hat s'_{i} \geq 1/i}$
\State $i'(i) = $ the index of $\hat s_i$ in $\hat s'$
\State Compute $A$ as the subset of infrequently observed arms $\{(i,j):i'(i) \leq \hat m, j = \ind{\hat q_{i} \leq \frac{1}{2}} \}$ with $|A| = \hat m$

\For{$(i,j) \in A$}
\For{$t \in 1,\ldots,h/2\hat m$}
\State Choose action $X_{i,t} = j$
\State Observe reward $r_t(X_{i,t}=j)$
\EndFor
\State Recompute $\hat \mu_{i,j} = \frac{2\hat m}{h} \sum_{t=1}^{ h/2\hat m} r_t(X_{i,t}=j)$ 
\EndFor
\end{algorithmic}
\end{algorithm}

\begin{theorem}\label{thm:simple-regret}
Define $m = \min\set{2 \leq i \leq N : q_i \geq 1/i}$.
Then the algorithm given in \cref{alg:simple} satisfies
\eq{
R_s(h) \in O\left(\sqrt{\frac{m}{h} \log \left(\frac{NT}{m}\right)}\right)\,.
}
\end{theorem}

\begin{lemma}\label{lem:conc1}
$\displaystyle \P{\left|\hat q_i - q_i\right| \geq \sqrt{\frac{6q_i}{h} \log \frac{2}{\delta}}} \leq \delta$.
\end{lemma}

\begin{proof}
Let $Z_t = \ind{X_{i,t} = 1} \in \set{0,1}$.
Then
\eq{
\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} Z_t\,.
}
Now $Z_1,\ldots,Z_{h/2}$ is an i.i.d.\ sequence of Bernoulli random variables with with mean $q_i$. The result follows from the Chernoff bound.
\end{proof}

\begin{lemma}\label{lem:conc2}
Let $\delta >0$ and $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$ and
define $\hat m = \min\set{2 \leq i \leq N : \hat s_{(i)} \geq 1/i}$.
If
\eq{
h \geq 24m \log\frac{4N}{\delta}\,,
}
then
\eq{
\P{\hat m \leq 2m} \geq 1 - \delta\,.
}
\end{lemma}

\begin{proof}
It is easy to see that the worst case occurs when 
\eq{
q_i = \begin{cases}
0 & \text{if } i < m \\
\frac{1}{m} & \text{otherwise}\,.
\end{cases}
}
Now by \cref{lem:conc1} we have with probability at least $1 - \delta$ that 
\eq{
(\forall i) \qquad \left| \hat q_i - q_i\right| 
&\leq \sqrt{\frac{6q_i}{h} \log \frac{2N}{\delta}} \\
&= \sqrt{\frac{6}{mh} \log\frac{2N}{\delta}}\,. 
}
By assumption 
\eq{
h \geq 24m \log \frac{4N}{\delta}\,.
}
Therefore with probability at least $1 - \delta$ we have
\eq{
(\forall i) \qquad \left|\hat q_i - q_i\right| \leq \frac{1}{2m}\,.
}
Note that for $i$ with $q_i = 0$ it is easy to see that $\hat q_i = 0$ (the variance is zero). The above guarantee that
for $i$ with $q_i = 1/m$ we have $\hat q_i \in [1/(2m), 3/(2m)]$ for all $i$ with probability at least $1 - \delta$,
which implies that $\hat s_i \geq 1/(2m)$ for all $i$ where $q_i = 1/m$.
Therefore $\hat m \leq 2m$ with probability at least $1 - \delta$. 
\end{proof}

\begin{lemma}\label{lem:m_est_small_enough}
Let $\delta >0$ and $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$ and
define $\hat m = \min\set{2 \leq i \leq N : \hat s_{(i)} \geq 1/i}$.
If
\eq{
h \geq 24m \log\frac{4N}{\delta}\,,
}
then
\eq{
\P{\hat m < \frac{2}{3}m} \leq \delta\,.
}
\end{lemma}

@@

To over-estimate $m$, we must identify as balanced some unbalanced arms (ie those with $i < m$. For all arms with $i < m$, by definition, $q_i < \frac{1}{m}$

Now by \cref{lem:conc1} we have with probability at least $1 - \delta$ that: \color{red}(this step not quite right - because I haven't said $q_i = 1/m$ as the worst case is that $q_i$ is infintesimally smaller than $1/m$. With smaller $q_i$ the bounds would be slightly tighter but not made up for by the decrease in the mean - ie we are more likely to over-estimate if $q_i = 1/m$ than if $q_i < 1/m$. Need a better way to say this ... ). Should probably just combine this with the other side into one lemma (even though we don't need this side for simple regret. \color{black}

\eq{
(\forall i) \qquad \left| \hat q_i - q_i\right| 
&\leq \sqrt{\frac{6q_i}{h} \log \frac{2N}{\delta}} \\
&= \sqrt{\frac{6}{mh} \log\frac{2N}{\delta}}\,. 
}

By assumption 
\eq{
h \geq 24m \log \frac{4N}{\delta}\,.
}
Therefore with probability at least $1 - \delta$ we have
\eq{
(\forall i < m) \qquad & \left|\hat q_i - q_i\right| \leq \frac{1}{2m}\,.\\
 & \hat{q}_i < \frac{3}{2}m \\
 \implies & \hat{m} > \frac{2}{3}m
}

@@

Now by \cref{lem:conc2}, 
\eq{
h \geq 24m \log \frac{4N}{\delta} \implies \P{\hat m > 2m} \leq \delta
}

@@

\subsubsection{Bounding the Bandit Regret}

@@

\begin{algorithm}[H]
\caption{Bandit Regret Algorithm}\label{alg:bandit}
\begin{algorithmic}[1]
\State {\bf Input:} $T, N$
\State $\delta = \frac{1}{T^{1/3}}$ 
\State $T_1 = 24Nlog\left(4N/\delta\right)$ 
\For{$t \in 1,T_1$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute $\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} X_{i,t}$
\State Compute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Compute $\hat{s'} = sorted(\hat{s}) : \hat{s'}_1 \leq \hat{s'_2} \leq ... \leq \hat{s'}_N$
\State Compute $\hat m = \min\set{2 \leq i \leq N : \hat s'_{i} \geq 1/i}$
\If {$\hat{m} > \frac{N^{3/2}}{\sqrt{T}}$}
\State Switch to the standard UCB algorithm with $K = 2N$ arms.
\Else
\State $h = T^{2/3}\hat{m}^{1/3}log(TK)^{1/3}$ 
\For{$t \in T_1,h/2$}
\State Do nothing and observe $X_{i,t}$ for $i \in \set{1,\ldots,N}$ and $r_t$
\EndFor
\State Compute for all $i \in \set{1,\ldots,N}$ and $j \in \set{0,1}$:
\eq{
\hat \mu_{i,j} = \frac{\sum_{t=1}^{h/2} \ind{X_{i,t} = j} r_t}{\sum_{t=1}^{h/2} \ind{X_{i,t} = j}}\,.
}
\State Recompute $\hat q_i = \frac{2}{h} \sum_{t=1}^{h/2} X_{i,t}$
\State Recompute $\hat s_i = \min\set{\hat q_i, 1 - \hat q_i}$
\State Recompute $\hat{s'} = sorted(\hat{s}) : \hat{s'}_1 \leq \hat{s'_2} \leq ... \leq \hat{s'}_N$
\State Recompute $\hat m = \min\set{2 \leq i \leq N : \hat s'_{i} \geq 1/i}$
\State $i'(i) = $ the index of $\hat s_i$ in $\hat s'$
\State Compute $A$ as the subset of infrequently observed arms $\{(i,j):i'(i) \leq \hat m, j = \ind{\hat q_{i} \leq \frac{1}{2}} \}$ with $|A| = \hat m$

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{comment}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{tikz} 
\usetikzlibrary{arrows,positioning} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}[1]{\mathbb E\left[#1\right]}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left(#1\right)}
\newcommand{\bigo}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\bigotilde}[1]{\tilde{\mathcal{O}}\left( #1 \right)}
\newcommand{\bigtheta}[1]{\Theta\left( #1 \right)}
\newcommand{\bigthetatilde}[1]{\tilde{\Theta}\left( #1 \right)}
\newcommand{\bigomega}[1]{\Omega\left( #1 \right)}

@@

\section{Proof of Theorem 1}
\begin{theorem}\label{thm:known_q_regret}
Define $m =   min\set{1 \leq i \leq N:q_{i+1} \geq \frac{1}{i}}$
Then Algorithm 1 satisfies
\eq{
R(T) \in \bigo{T^{2/3}m^{1/3}log(KT)^{1/3}}\,.
}
\end{theorem}

Let $A = \set{(i,j) : i \leq m, j = 1}$ be the set of infrequently observed arms

For the frequently observed arms, $(i,j) \notin A$ we have:

\eqn {
\hat{\mu}_{i,j} = \frac{2}{h}\frac{\sum_{t=1}^{h/2}\ind{X_{i,t}=j}r_t}{q_i^j(1-q_i)^{1-j}}
}

@@

The algorithm explicitly plays each of the infrequently observed arms, $(i,j) \in A$, $\frac{h}{2m}$ times. So:

@@

\section{Proof of Theorem 2}

\begin{theorem}\label{thm:simple-regret}
Define $m = \min\set{2 \leq i \leq N : q_i \geq 1/i}$.
Then Algorithm 2 satisfies
\eq{
R^{simple}(T) \in O\left(\sqrt{\frac{m}{T} \log \left(\frac{NT}{m}\right)}\right)\,.
}
\end{theorem}

\begin{lemma}\label{lem:conc1}
$\displaystyle \P{\left|\hat q_i - q_i\right| \geq \sqrt{\frac{6q_i}{T} \log \frac{2}{\delta}}} \leq \delta$.
\end{lemma}

\begin{proof}
Let $Z_t = \ind{X_{i,t} = 1} \in \set{0,1}$.
Then
\eq{
\hat q_i = \frac{2}{T} \sum_{t=1}^{T/2} Z_t\,.
}
Now $Z_1,\ldots,Z_{T/2}$ is an i.i.d.\ sequence of Bernoulli random variables with with mean $q_i$. The result follows from the Chernoff bound.
\end{proof}
\begin{lemma}\label{lem:m_est}
Let $\delta >0.$ 
If $h \geq 24m \log\frac{4N}{\delta}$

then
\eq{
\P{\hat m < \frac{2}{3}m} \leq \delta\ \text{ and } \P{\hat m > 2m} \leq \delta\,.
}
\end{lemma}

@@

Let $\boldsymbol{q}^b$ and $\boldsymbol{q}^{ub}$ and be the maximally balanced and unbalanced $\boldsymbol{q}$ for a given $m$ respectively.

\eq{
q^{ub}_i = \begin{cases}
0 & \text{if } i \leq m \\
\frac{1}{m} & \text{otherwise}\,.
\end{cases}\\
q^b_i < \begin{cases}
\frac{1}{m} & \text{if } i \leq m \\
1 & \text{otherwise}\,.
\end{cases}  
}
For $\hat{m}$ to over-estimate $m$, we must identify some balanced arms as unbalanced. For $\hat{m}$ to under-estimate $m$, we must identify some unbalanced arms as balanced.
\eq{
\P{\hat m > 2m} \leq \P{\hat m > 2m|\boldsymbol{q} = \boldsymbol{q}^{ub}}\\
\P{\hat m < \frac{2}{3}m} \leq \P{\hat m < \frac{2}{3}m|\boldsymbol{q} = \boldsymbol{q}^{b}}
}

Given $\boldsymbol{q} = \boldsymbol{q}^{ub}$, we have by \cref{lem:conc1}, with probability at least $1 - \delta$ that:

\eq{
& \left| \hat q_i - q_i\right| 
 \leq \begin{cases}
0 & \text{if } i \leq m\\
\sqrt{\frac{6}{mh} \log\frac{2}{\delta}} & \text{otherwise} \\
\end{cases}\\
\implies &
\begin{cases}
(\forall i \leq m) \qquad \left| \hat q_i - 0\right| = 0 \\
(\forall i > m) \qquad \left| \hat q_i - \frac{1}{m}\right| \leq \frac{1}{2m} \; \text{, taking the union bound and assuming } h \geq 24m \log\frac{4N}{\delta}\\
\end{cases}\\
\implies &
\begin{cases}
(\forall i \leq m) \qquad  \hat q_i  = 0 \\
(\forall i > m) \qquad \hat q_i \in [\frac{1}{2m}, \frac{3}{2m}]\\
\end{cases}\\
\implies & \hat{m} \leq 2m
}

Given $\boldsymbol{q} = \boldsymbol{q}^{b}$, we have by \cref{lem:conc1}, with probability at least $1 - \delta$ that:

\eq{
& \left| \hat q_i - q_i\right| 
 \leq \sqrt{\frac{6}{mh} \log\frac{2}{\delta}} \qquad \text{if } i \leq m \\
\implies &
(\forall i \leq m) \qquad  \hat q_i \leq \frac{3}{2m} \\
\implies & \hat{m} \geq \frac{2m}{3}
}

@@

Now by \cref{lem:m_est}, 
\eq{
h \geq 24m \log \frac{4N}{\delta} \implies \P{\hat m > 2m} \leq \delta
}

@@

\section{Proof of Theorem 3}

\begin{theorem}\label{thm:unknown_q_regret}
Define $m =   min\set{1 \leq i \leq N:q_{i+1} \geq \frac{1}{i}}$
Then Algorithm 3 satisfies
\eq{
R(T) \in \bigo{T^{2/3}m^{1/3}log(KT)^{1/3}}\,.
}
\end{theorem}

@@
